Joint estimation of phylogenies and ancestral states is not consistent

### 2017-12-07 joint_inf.tex
clarify

%EM Question: it seems to me that even in the non-white region, the BL inferences aren't correct. Is that true? We should say so if. I don't want to make extra work for you, but a figure in which you take the difference between the correct value and the inferred value might be nice.
%das: it seemed like the estimates were decent to me, but I can do this super easily (output is pickled). Though are you talking about another plot or just a check that we can add a sentence like "even in cases where \hat{w} is not one w is overestimated/poorly estimated."
%EM I'm pretty sure I can see that there's bias in the estimates even in the non-white area. A plot of the difference between the two would be cool. Can we zoom into the region with x* and y* greater than 0.6 so that the scale doesn't get thrown off by big differences where x* and y* are zero?
Marginal inference performs as expected, where $\hat{w}$ is equal to $y^*$ regardless of the value of $x^*$ (Fig.~\ref{fig:bl-general-marginal}) when optimizing \eqref{eq:marginal_likelihood} using the same procedure.
For joint inference, the estimates for $\hat{w}$ when $x^*$ and $y^*$ are both large look reasonable, with $\hat{w}$ increasing as $y^*$ increases, though Fig.~\ref{fig:bl-general-bias} shows there is a systematic positive bias in this procedure.
% Here we obtain the correct $\hat{w}\rightarrow y^*$.
To understand the quality of each fit, we report the range of $\hat{w}-y^*$ where $\hat{w}\neq 1$.
For joint inference, the errors range from $[-7\times 10^{-3}, 8\times 10^{-2}]$ and for marginal inference, $[-8\times 10^{-8}, 5\times 10^{-7}]$ showing that, even in cases where joint inference does not estimate $\hat{w}$ to be exactly one, it still fails to achieve a low error from truth.

@@

While we have shown inconsistency in both topology and branch parameter estimation, there is substantial scope for future work to make these results more precise and more general.
The techniques used to obtain upper and lower bounds for the likelihoods in the topology estimation case provide relatively loose bounds, though how loose they are remain unknown without either further analysis or verification through simulation.
Similarly, for the general case in estimating branch lengths, we were only able to provide a conservative region of overestimation, and the unusual shape we observe via numerical optimization (Fig.~\ref{fig:bl-general-inconsistency}) begs further investigation.
Empirical validation shows that the general case is not unlike the restricted case.
All of these results hold only for a simple binary symmetric model on four-taxon trees, and extensive simulation is necessary to understand how these results extend to more complicated general cases.

>>>>

%EM Question: it seems to me that even in the non-white region, the BL inferences aren't correct. Is that true? We should say so if. I don't want to make extra work for you, but a figure in which you take the difference between the correct value and the inferred value might be nice.
%das: it seemed like the estimates were decent to me, but I can do this super easily (output is pickled). Though are you talking about another plot or just a check that we can add a sentence like "even in cases where \hat{w} is not one w is overestimated/poorly estimated."
%EM I'm pretty sure I can see that there's bias in the estimates even in the non-white area. A plot of the difference between the two would be cool. Can we zoom into the region with x* and y* greater than 0.6 so that the scale doesn't get thrown off by big differences where x* and y* are zero?
Marginal inference performs as expected, where $\hat{w}$ is equal to $y^*$ regardless of the value of $x^*$ (Fig.~\ref{fig:bl-general-marginal}) when optimizing \eqref{eq:marginal_likelihood} using the same procedure.
For joint inference, the estimates for $\hat{w}$ when $x^*$ and $y^*$ are both large look reasonable, with $\hat{w}$ increasing as $y^*$ increases, though Fig.~\ref{fig:bl-general-bias} shows there is a systematic positive bias in this procedure when $\hat{w}$ is not estimated to be 1.
% Here we obtain the correct $\hat{w}\rightarrow y^*$.
To understand the quality of each fit, we report the range of $\hat{w}-y^*$ where $\hat{w}\neq 1$.
For joint inference, the errors range from $[-7\times 10^{-3}, 8\times 10^{-2}]$ and for marginal inference, $[-8\times 10^{-8}, 5\times 10^{-7}]$ showing that, even in cases where joint inference does not estimate $\hat{w}$ to be exactly one, it still fails to achieve a low error from truth.

@@

While we have shown inconsistency in both topology and branch parameter estimation, there is substantial scope for future work to make these results more precise and more general.
The techniques used to obtain upper and lower bounds for the likelihoods in the topology estimation case provide relatively loose bounds, though how loose they are remain unknown without either further analysis or verification through simulation.
Similarly, for the general case in estimating branch lengths, we were only able to provide a conservative region of overestimation, and the unusual shape we observe via numerical optimization (Fig.~\ref{fig:bl-general-inconsistency}) begs further investigation.
Empirical validation shows that the general case is not unlike the restricted case.
All of these results hold only for a simple binary symmetric model on four-taxon trees, and extensive simulation is necessary to understand how these results extend to more complicated general cases.
%EM Any intuition of if the effect will go away on larger trees? What about more complex models?
